### 2016-11-18 Manuscript/SW_early_life_telomeres.Rmd
Eds

**Early-life conditions, telomere dynamics and survival: insights from a long-term island bird study**

@@

3. We found that telomere length decreases with cross-sectional and longitudinal measures of age, and that telomere length decreases most rapidly early in life. However, both cross-sectional and longitudinal data suggested that telomere length increases occurs in some individuals, and using a large number of repeated measurements we show statistically that this increase is unlikely to be explained by qPCR measurement error.

@@

#####Study species and sampling
The Seychelles warbler is a small (~15 g), insectivorous passerine bird with a mean life expectancy of 5.5 years at fledging [@Hammers2013]. The population of *ca*. 320 birds on Cousin Island (04'20'S, 55'40'E) has been intensively studied since 1986 [@Richardson2003b; @Spurgin2014a]. This species' main breeding season runs from June-August, when the breeding females on each of the *ca*. 115 territories lay one or, rarely, two or three eggs [@Komdeur1991]. As a result of this low reproductive output, combined with higher mortality in first-year birds [39%; @Brouwer2006], cohort sizes in the Seychelles warbler are small (<50).

@@

Because we observed apparent telomere lengthening in our data, we used the entire Seychelles warbler dataset to test whether this lengthening could be explained by measurement error. Using all birds (adults and juveniles) with at least two telomere measurement, we calculated $\Delta$RTL for each pair of samples (for example, for a bird with three longitudinal measurements throughout their life we calculated the change in RTL between the first and second sample and between the second and third sample). We then compared these longitudinal changes in telomere length to sampling error, using all samples where we had at least two repeat telomere measurements for the same sample (i.e. completely seperate reactions run on separate plates: N = `r nrow(subset(ddL,Group == 'Within samples'))` measurements from `r length(unique(subset(ddL,Group == 'Within samples')$BirdID))` samples). We calculated $\Delta$RTL between pairs of repeat measurements within samples in exactly the same way as for across samples, except that repeat measurements were ordered by the date at the qPCR was run rather than by age.

We used LMMs to explore how variation in early-life environmental and social conditions influenced telomere length within cohorts. We created a full model with RTL as a response variable, alongside the following explanatory variables: age, tarsus length, body mass, sex, annual food availability, population density (an annual measure estimated from the summer breeding census) and the number of helping and non-helping subordinate birds present in the natal territory. As random effects we included birth year and individual identity. We report model estimates and confidence intervals for all effects included in the full model. As our analysis is exploratory, and because multicollinearity between variables in our full model may have obscured real effects, we also performed model averaging, using the MuMIn package in R [@Barton2012]. Model selection was performed using the full model described above and a top model set defined, containing all models with AICc $\leq$ 6 compared to the best supported model [@Burnham2011]. We report model-averaged coefficients, confidence intervals and 'relative importance', which reflects the relative weights of each predictor variable across the top model set. For individuals with longitudinal data (*n* = `r nrow(dd3)`) we repeated the above analyses of telomere dynamics, replacing telomere length with $\Delta$RTL as the response variable, and excluding the plate ID random effect (as each measurement of $\Delta$RTL was based on two or more measurements, and so run on multiple plates).

@@

Including birth year as a random effect in a LMM of RTL and (log) age resulted in a significantly improved model to one with only individual ID as a random effect (P `r round3(anova(RTL_LogAge,RTL_LogAge_LY)$"Pr(>Chisq)"[2])`; Fig. 2A). Allowing the slope of the age effect to vary among birth years resulted in a slightly significant improvement on one in which only the intercept was allowed to vary (P = `r round3(anova(RTL_LogAge_LY,RTL_LogAge_AgeLY)$"Pr(>Chisq)"[2])`; Fig. 2B). However, a model including catch year and individual ID as random effects was significantly better than a model indluding birth year and individual ID (P `r round3(anova(RTL_LogAge_LY,RTL_LogAge_CY)$"Pr(>Chisq)"[2])`) suggesting that variation among catch years was greater than among cohorts (Fig. 2C), and a model that allowed the slope of the age effect to vary among catch years was better still (P = `r round3(anova(RTL_LogAge_CY,RTL_LogAge_AgeCY)$"Pr(>Chisq)"[2])`; Fig. 2D). Using longitudinal data and $\Delta$Age as an explanatory variable we found a similar pattern - a model including catch year as a random effect was better than that did not allow for temporal variation in RTL (P = `r round3(anova(RTL_DeltaAge,RTL_DeltaAge_CY)$"Pr(>Chisq)"[2])`), and better than one with birth year as a random effect (P `r round3(anova(RTL_DeltaAge_LY,RTL_DeltaAge_CY)$"Pr(>Chisq)"[2])`). However a model allowing the slope of the $\Delta$Age effect was not better than one in which only the intercept varied (P = `r round3(anova(RTL_DeltaAge_CY,RTL_DeltaAge_AgeCY)$"Pr(>Chisq)"[2])`).

Within individuals, longitudinal data showed that RTL measurements taken from different timepoints were weakly, but significantly, positvely correlated (R^2^ = `r round3(summary(lm(RTL1~RTL,data = dd3))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL,data = dd3))$coefficients[2,3])`; P = `r round3(summary(lm(RTL1~RTL,data = dd3))$coefficients[2,4])`; Fig. 1B). Although both or cross-sectional and longitudinal data indicated a general trend of telomere shortening with age, we found that `r round((nrow(subset(dd3,DeltaRTL>0))/nrow(dd3))*100,0)`% of our $\Delta$RTL measurments were positive, and thus indicative of telomere lengthening. To test whether the apparent telomere lengthening in our dataset could be explained by measurement error, we compared variance in telomere length among repeat measurements of samples to the variance observed among different samples of the same individual (see methods for details). We found significantly higher variance in telomere length over individual lifetimes compared to among sample replicates (Levene's test: F = `r round2(leveneTest(DeltaRTL~Group,data=ddL)['F value'][1,1])`; P `r round3(leveneTest(DeltaRTL~Group,data=ddL)['Pr(>F)'][1,1])`; Fig. 1C). Importantly, splitting the longitudinal data into instances of shortening ($\Delta$RTL < 0) and lenghtning ($\Delta$RTL > 0) revealed that not only did we observe significantly more shortening within individuals compared to within samples (Wilcoxon test: W = `r round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$statistic,0)`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$p.value)`), but also significantly more lengthening (W = `r as.character(round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$statistic,0))`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$p.value)`).

#####Spatiotemporal variation in telomere dynamics and the environment
A linear mixed model of RTL in relation to social and ecological environmental variables revealed age, tarsus length, sex and annual variation in food availability were significantly related to RTL (Table 1). The model averaging approach yielded qualitatively identical results to the full LMM, with the same explanatory variables 'significant' in terms of being retained in top models, and having model-averaged confidence intervals not overlapping zero (Table S1; Fig. S1). One interesting finding from the model selection was that sex only appeared in top models where tarsus length was also present (Table S1). In accordance with this, when tarsus length was removed from the full model sex was no longer significant (estimate = `r round3(fixef(RTL_Full_NoTarsus)['SexMales'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Full_NoTarsus)['SexMales',],lessthan = F),collapse = ', ')`), and a sex x tarsus interaction was significant when included (estimate = `r round3(fixef(RTL_Full_SexTarsus)['SexMales:Tarsus'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Full_SexTarsus)['SexMales:Tarsus',],lessthan = F),collapse = ', ')`); RTL increased with tarsus length in males, but decreased in females.

No social or ecological environmental variables were significant in a model of longitudinal data with $\Delta$RTL as the response variable (Table 2). The model averaging approach suggested that $\Delta$RTL was positively related to age (i.e. telomere shortening increased with age), and negatively associated with annual density (i.e. more shortening in years with higher population density; Table S2; Fig. S2).

###Discussion
Here we use a large, long-term, multi-cohort dataset to assess the relationships between the spatiotemporal variation in the ecological environment and lifelong telomere dynamics in a closed population of Seychelles warblers. We found that, as in other studies, telomeres shorten over time, and that telomere loss is greatest in early life. However, bouts of apparent telomere lengthening occur, and the extent of this lengthening cannot be explained solely by measurement error. We then found that telomere dynamics vary among sexes, and over time, with this temporal variation is related to fluctuations in food availability.

@@

###Acknowledgements
We thank Nature Seychelles for facilitating the long-term Seychelles warbler project. The Seychelles Bureau of Standards and Department of Environment gave permission for sampling and fieldwork. Emma Barrett laid the foundations for this study, generating the original telomere qPCR protocol. We thank everyone who has helped in the field, and the Seychelles warbler research group for discussions. This work was funded by two Natural Environment Research Council (NERC) grants to DSR (NE/F02083X/1 and NE/K005502/1). LGS was also funded by fellowships from the University of Oxford and the BBSRC, and HLD by a NERC fellowship.

>>>>

**Spatiotemporal variation in lifelong telomere dynamics in the Seychelles warbler**

@@

3. We found that telomere length decreases with cross-sectional and longitudinal measures of age, and that telomere length decreases most rapidly early in life. However, both cross-sectional and longitudinal data suggested that telomere length increases occurs in some individuals for short periods, and using a large number of repeated measurements we show statistically that this increase cannot be explained solely by qPCR measurement error.

@@

#####Study species and sampling
The Seychelles warbler is a small (~15 g), insectivorous passerine bird with a mean life expectancy of 5.5 years at fledging [@Hammers2013]. The population of *ca*. 320 birds on Cousin Island (04'20'S, 55'40'E) has been intensively studied since 1986 [@Richardson2003b; @Spurgin2014a]. This species' main breeding season runs from June-August, when the breeding females on each of the *ca*. 115 territories lay one or, rarely, two or three eggs [@Komdeur1991]. As a result of this low reproductive output, combined with higher mortality in first-year birds [39%; @Brouwer2006], cohort sizes in the Seychelles warbler are small (< 50).

@@

Because we observed apparent telomere lengthening in our data, we used the entire Seychelles warbler dataset to test whether this lengthening could be explained by measurement error. Using all birds (adults and juveniles) with at least two telomere measurement, we calculated $\Delta$RTL for each pair of samples (for example, for a bird with three longitudinal measurements throughout their life we calculated the change in RTL between the first and second sample and between the second and third sample). We then compared these longitudinal changes in telomere length to sampling error, using all samples where we had at least two repeat telomere measurements for the same sample (i.e. completely seperate reactions run on separate plates: N = `r nrow(subset(ddL,Group == 'Within samples'))` measurements from `r length(unique(subset(ddL,Group == 'Within samples')$BirdID))` samples). We calculated $\Delta$RTL between pairs of repeat measurements within samples in exactly the same way as for across samples, except that repeat measurements were ordered by the date at the qPCR was run rather than by age. We first compared the variance in $\Delta$RTL within and among individuals using a Levene's test. Then, to separately test whether the extent of telomere shortening and lenghtening within individuals was greater than expected by measurement error, we split $\Delta$RTL measurements into groups in which shortening ($\Delta$RTL < 0) and lengthening $\Delta$RTL ($\Delta$RTL > 0) occurred, and tested whether $\Delta$RTL among samples was significantly different between within-sample and across-sample measurements, using Wilcoxon tests.

We used LMMs to test how variation in early-life environmental and social conditions influenced telomere length within cohorts. We created a full model with RTL as a response variable, alongside the following explanatory variables: age, tarsus length, body mass, sex, annual food availability, population density (an annual measure estimated from the summer breeding census) and the number of helping and non-helping subordinate birds present in the natal territory. As random effects we included birth year and individual identity. We report model estimates and confidence intervals for all effects included in the full model. As our analysis is exploratory, and because multicollinearity between variables in our full model may have obscured real effects, we also performed model averaging, using the MuMIn package in R [@Barton2012]. Model selection was performed using the full model described above and a top model set defined, containing all models with AICc $\leq$ 6 compared to the best supported model [@Burnham2011]. We report model-averaged coefficients, confidence intervals and 'relative importance', which reflects the relative weights of each predictor variable across the top model set. For individuals with longitudinal data (*n* = `r nrow(dd3)`) we repeated the above analyses of telomere dynamics, replacing telomere length with $\Delta$RTL as the response variable, and excluding the plate ID random effect (as each measurement of $\Delta$RTL was based on two or more measurements, and so run on multiple plates).

@@

Including birth year as a random effect in a LMM of RTL and (log) age resulted in a significantly improved model to one with only individual ID ans plate ID as random effects (P `r round3(anova(RTL_LogAge,RTL_LogAge_LY)$"Pr(>Chisq)"[2])`; Fig. 2A). Allowing the slope of the age effect to vary among birth years resulted in a slightly significant improvement on one in which only the intercept was allowed to vary (P = `r round3(anova(RTL_LogAge_LY,RTL_LogAge_AgeLY)$"Pr(>Chisq)"[2])`; Fig. 2B). However, a model including catch year and individual ID as random effects was significantly better than a model indluding birth year and individual ID (P `r round3(anova(RTL_LogAge_LY,RTL_LogAge_CY)$"Pr(>Chisq)"[2])`; Fig 2C) suggesting that variation among catch years was greater than among cohorts (Fig. 2C), and a model that allowed the slope of the age effect to vary among catch years was better still (P = `r round3(anova(RTL_LogAge_CY,RTL_LogAge_AgeCY)$"Pr(>Chisq)"[2])`; Fig. 2D). Using longitudinal data and $\Delta$Age as an explanatory variable we found a similar pattern - a model including catch year as a random effect was better than that did not allow for temporal variation in RTL (P = `r round3(anova(RTL_DeltaAge,RTL_DeltaAge_CY)$"Pr(>Chisq)"[2])`), and better than one with birth year as a random effect (P `r round3(anova(RTL_DeltaAge_LY,RTL_DeltaAge_CY)$"Pr(>Chisq)"[2])`). However a model allowing the slope of the $\Delta$Age effect was not better than one in which only the intercept varied (P = `r round3(anova(RTL_DeltaAge_CY,RTL_DeltaAge_AgeCY)$"Pr(>Chisq)"[2])`).

Within individuals, longitudinal data showed that RTL measurements taken from different timepoints were weakly, but significantly, positvely correlated (R^2^ = `r round3(summary(lm(RTL1~RTL,data = dd3))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL,data = dd3))$coefficients[2,3])`; P = `r round3(summary(lm(RTL1~RTL,data = dd3))$coefficients[2,4])`; Fig. 3A). Although both or cross-sectional and longitudinal data indicated a general trend of telomere shortening with age, we found that `r round((nrow(subset(dd3,DeltaRTL>0))/nrow(dd3))*100,0)`% of our $\Delta$RTL measurments were positive, and thus indicative of telomere lengthening. To test whether the apparent telomere lengthening in our dataset could be explained by measurement error, we compared variance in telomere length among repeat measurements of samples to the variance observed among different samples of the same individual (see methods for details). We found significantly higher variance in telomere length over individual lifetimes compared to among sample replicates (Levene's test: F = `r round2(leveneTest(DeltaRTL~Group,data=ddL)['F value'][1,1])`; P `r round3(leveneTest(DeltaRTL~Group,data=ddL)['Pr(>F)'][1,1])`; Fig. 3B). Importantly, splitting the longitudinal data into instances of shortening ($\Delta$RTL < 0) and lenghtning ($\Delta$RTL > 0) revealed that not only did we observe significantly more shortening within individuals compared to within samples (Wilcoxon test: W = `r round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$statistic,0)`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$p.value)`), but also significantly more lengthening (W = `r as.character(round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$statistic,0))`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$p.value)`; Fig. 3B).

#####Spatiotemporal variation in telomere dynamics and the environment
A LMM revealed that age, tarsus length, sex and annual variation in food availability were significantly related to RTL (Table 1). Tarsus length was negatively related to RTL and males had longer telomeres than females (Table 1; Fig. 4A), while insect availability was positively related to RTL (Table 1; Fig. 4B). The model averaging approach yielded qualitatively identical results to the full LMM, with the same explanatory variables 'significant' in terms of being retained in top models, and having model-averaged confidence intervals not overlapping zero (Table S1; Fig. S1). One interesting finding from the model selection was that sex only appeared in top models where tarsus length was also present (Table S1). In accordance with this, when tarsus length was removed from the full model sex was no longer significant (estimate = `r round3(fixef(RTL_Full_NoTarsus)['SexMales'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Full_NoTarsus)['SexMales',],lessthan = F),collapse = ', ')`), and a sex x tarsus interaction was significant when included (estimate = `r round3(fixef(RTL_Full_SexTarsus)['SexMales:Tarsus'],lessthan = F)`, CIs = `r paste(round3(CIwald(RTL_Full_SexTarsus)['SexMales:Tarsus',],lessthan = F),collapse = ', ')`); RTL increased with tarsus length in males, but decreased in females (Fig. 4A).

No social or ecological environmental variables were significant in a model of longitudinal data with $\Delta$RTL as the response variable (Table 2). The model averaging approach suggested that $\Delta$RTL was positively related to age (i.e. telomere shortening increased with age), and negatively associated with annual density (i.e. more shortening in years with higher population density; Table S2; Figs S2; S3).

###Discussion
Here we use a long-term, multi-cohort dataset to assess the relationships between the spatiotemporal variation in the ecological environment and lifelong telomere dynamics in a closed population of Seychelles warblers. We found that telomeres shorten over time, and that telomere loss is greatest in early life. However, bouts of apparent telomere lengthening occur, and the extent of this lengthening cannot be explained solely by measurement error. We then found that telomere dynamics vary among sexes, and over time, with this temporal variation is related to fluctuations in food availability.

@@

###Acknowledgements
We thank Nature Seychelles for facilitating the long-term Seychelles warbler project. The Seychelles Bureau of Standards and Department of Environment gave permission for sampling and fieldwork. Emma Barrett laid the foundations for this study, generating the original telomere qPCR protocol. We thank everyone who has helped in the field, and the Seychelles warbler research group for discussions. This work was funded by two Natural Environment Research Council (NERC) grants to DSR (NE/F02083X/1 and NE/K005502/1). LGS was also funded by a fellowship from the BBSRC, and HLD by a NERC fellowship.

### 2016-11-18 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
Eds

Fig2A <- ggplot(dd3,
       aes(x = RTL,
           y = RTL1)) +
  ylab(expression('RTL ('*italic(t)*' + 1)')) + 
  xlab(expression('RTL ('*italic(t)*')')) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c('none'),
        legend.title = element_text(size = 12),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black')) +
    geom_point(col = 'grey')+
  geom_abline(slope = 1,intercept = 0,lty = 2)
  annotate('text',x = 0,y = 3,label = 'B')

Fig2B <- ggplot(ddL,aes(col=Group,fill = Group,y = ..scaled..,x = DeltaRTL))+
    geom_density(alpha = 0.2)+
  theme_classic()+
  ylab ('Scaled density')+
  xlab(expression(Delta*'RTL'))+
    theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none',
        legend.title = element_blank(),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black')) +
  scale_colour_manual(values = c('darkgrey','gold'))+
  scale_fill_manual(values = c('darkgrey','gold'))+
  geom_vline(xintercept = 0,lty = 2)
  annotate('text',x = -2,y = 1,label = 'C')

@@

Fig3A <- transform(juv,Cohort = factor(CatchYear)) %>%
  subset(Agemonths>(1/12)) %>%
  ggplot(aes(y = RTL,x = Cohort))+
  xlab('Cohort') +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12,angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'))
  annotate('text',x = 0,y = 1.4,label = 'B')

Fig3B <- transform(dd,Cohort = factor(CatchYear)) %>%
  ddply(.(Cohort),
        mutate,
        nc = length(which(Agemonths ==1/12))) %>%
  subset(nc > 11) %>%
  ggplot(aes(y = RTL,x = Agemonths, col = Cohort))+
  xlab('Age (years)') +
  geom_smooth(method = 'lm', formula = y~log(x),se = F) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'))
  annotate('text',x = 0,y = 1.4,label = 'B')

>>>>

Fig2A <- transform(juv,Cohort = factor(CatchYear)) %>%
  subset(Agemonths>(1/12)) %>%
  ggplot(aes(y = RTL,x = Cohort))+
  xlab('Cohort') +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12,angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'))
  annotate('text',x = 0,y = 1.4,label = 'B')

Fig2B <- transform(dd,Cohort = factor(CatchYear)) %>%
  ddply(.(Cohort),
        mutate,
        nc = length(which(Agemonths ==1/12))) %>%
  subset(nc > 11) %>%
  ggplot(aes(y = RTL,x = Agemonths, col = Cohort))+
  xlab('Age (years)') +
  geom_smooth(method = 'lm', formula = y~log(x),se = F) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'))
  annotate('text',x = 0,y = 1.4,label = 'B')

@@

Fig3A <- ggplot(dd3,
       aes(x = RTL,
           y = RTL1)) +
  ylab(expression('RTL ('*italic(t)*' + 1)')) + 
  xlab(expression('RTL ('*italic(t)*')')) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c('none'),
        legend.title = element_text(size = 12),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black')) +
    geom_point(col = 'grey')+
  geom_abline(slope = 1,intercept = 0,lty = 2)
  annotate('text',x = 0,y = 3,label = 'B')

Fig3B <- ggplot(ddL,aes(col=Group,fill = Group,y = ..scaled..,x = DeltaRTL))+
    geom_density(alpha = 0.2)+
  theme_classic()+
  ylab ('Scaled density')+
  xlab(expression(Delta*'RTL'))+
    theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none',
        legend.title = element_blank(),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black')) +
  scale_colour_manual(values = c('darkgrey','gold'))+
  scale_fill_manual(values = c('darkgrey','gold'))+
  geom_vline(xintercept = 0,lty = 2)
  annotate('text',x = -2,y = 1,label = 'C')
